---
title: Proving the Central Limit Theorem
author: Vanshaj S. • Aug. 3, 2018
output: 
  ioslides_presentation:
    widescreen: true
    css: ios.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning=F, message=F, error=F, echo=F)
```

## Preparations
```{r, echo = T, results='hide'}
library(plotly)
library(manhattanly)

set.seed(81706)
```


## What is the CLT?
> In probability theory, the central limit theorem (CLT) establishes that, in some situations, when independent random variables are added, their properly normalized sum tends toward a normal distribution (informally a "bell curve") even if the original variables themselves are not normally distributed. The theorem is a key concept in probability theory because it implies that probabilistic and statistical methods that work for normal distributions can be applicable to many problems involving other types of distributions.

Source: Wikipedia

## Sample Sizes
Another aspect of the Central Limit Theorem is that as you increase the number of samples taken from a non-normal distribution, the distribution of means tends towards normalcy. In this presentation, we're going to explore exactly that.

## ns = 1
We're going to use a uniform distribution for our model, where n, the sample size, is 100. ns represents the number of samples taken.

```{r}
x <- runif(100)
fit <- density(x)

plot_ly(x = x, type = "histogram", name = "Data") %>% 
    add_trace(x = fit$x, y = fit$y, type = "scatter", mode = "lines", fill = "tozeroy", yaxis = "y2", name = "Density") %>%
    layout(yaxis2 = list(overlaying = "y", side = "right"))
```

## ns = 1

```{r}
qqnorm(x)
```

Not linear at all. Let's see how taking multiple samples changes this.

## ns = 10
Now let's take 10 samples on new uniform distributions and see what shape the means take.

```{r}
x <- replicate(10, runif(100))
fit <- density(colMeans(x))

plot_ly(x = colMeans(x), type = "histogram", name = "Data") %>% 
    add_trace(x = fit$x, y = fit$y, type = "scatter", mode = "lines", fill = "tozeroy", yaxis = "y2", name = "Density") %>%
    layout(yaxis2 = list(overlaying = "y", side = "right"))
```

## ns = 10

```{r}
qqnorm(colMeans(x))
```

Not a very linear quantile plot; the data isn't **too** normal.

## ns = 30
Let's ramp up to ns = 30, a number that I had learned in my high school AP Statistics class.

```{r}
x <- replicate(30, runif(100))
fit <- density(colMeans(x))

plot_ly(x = colMeans(x), type = "histogram", name = "Data") %>% 
    add_trace(x = fit$x, y = fit$y, type = "scatter", mode = "lines", fill = "tozeroy", yaxis = "y2", name = "Density") %>%
    layout(yaxis2 = list(overlaying = "y", side = "right"))
```

## ns = 30

```{r}
qqnorm(colMeans(x))
```

Looks slightly more linear and normal, but I'm not satisfied.

## ns = 50
Let's go a little higher.

```{r}
x <- replicate(50, runif(100))
fit <- density(colMeans(x))

plot_ly(x = colMeans(x), type = "histogram", name = "Data") %>% 
    add_trace(x = fit$x, y = fit$y, type = "scatter", mode = "lines", fill = "tozeroy", yaxis = "y2", name = "Density") %>%
    layout(yaxis2 = list(overlaying = "y", side = "right"))
```

## ns = 50

```{r}
qqnorm(colMeans(x))
```

This looks rather linear/normal, actually.

## ns = 100
Let's double it.

```{r}
x <- replicate(100, runif(100))
fit <- density(colMeans(x))

plot_ly(x = colMeans(x), type = "histogram", name = "Data") %>% 
    add_trace(x = fit$x, y = fit$y, type = "scatter", mode = "lines", fill = "tozeroy", yaxis = "y2", name = "Density") %>%
    layout(yaxis2 = list(overlaying = "y", side = "right"))
```

## ns = 100

```{r}
qqnorm(colMeans(x))
```

That's a very normal set of data, except for the weird valley in the middle of the histogram.

## ns = 500
If 100 fetches us good results, 500 should be nearly perfect.

```{r}
x <- replicate(500, runif(100))
fit <- density(colMeans(x))

plot_ly(x = colMeans(x), type = "histogram", name = "Data") %>% 
    add_trace(x = fit$x, y = fit$y, type = "scatter", mode = "lines", fill = "tozeroy", yaxis = "y2", name = "Density") %>%
    layout(yaxis2 = list(overlaying = "y", side = "right"))
```

## ns = 500

```{r}
qqnorm(colMeans(x))
```

Almost looks like the chart **is** a line.

## ns = 1000
Finally, let's see if 1000 samples give us a perfect distribution.

```{r}
x <- replicate(1000, runif(100))
fit <- density(colMeans(x))

plot_ly(x = colMeans(x), type = "histogram", name = "Data") %>% 
    add_trace(x = fit$x, y = fit$y, type = "scatter", mode = "lines", fill = "tozeroy", yaxis = "y2", name = "Density") %>%
    layout(yaxis2 = list(overlaying = "y", side = "right"))
```

## ns = 1000

```{r}
qqnorm(colMeans(x))
```

This is, essentially, as normal as a non-infinite dataset can probably get.
